|

### 2021-10-16 poster/GSPA_Abstract.Rmd
last changes prior to submission GSPA + ORA

We sampled 1,498 participants from Prolific (native English speakers from the US and UK, over 18 years old). As preregistered, we regressed all outcomes on prior beliefs and condition (with the single estimate condition as the reference category). Compared to the single estimate condition, inconsistent crowd estimates decreased posterior beliefs (*b* = -22.80, *p* < .001, Cohen's *d* = -0.72), ratings of credibility (*b* = -6.54, *p* < .001, *d* = -0.21), and confidence in the parameter estimate (*b* = -9.09, *p* < .001, *d* = -0.26), and increased ratings of bias (*b* = 6.68, *p* < .001, *d* = 0.18), error (*b* = 5.89, *p* < .001, *d* = 0.18), and discretion (*b* = 9.22, *p* < .001, *d* = 0.29). Compared to the single estimate condition, consistent crowd estimates decreased posterior beliefs (*b* = -5.71, *p* < .001, *d* = -0.18) and increased ratings of error (*b* = 3.59, *p* = .016, *d* = 0.11) and discretion (*b* = 5.14, *p* < .001, *d* = 0.16); we found no significant effects on ratings of credibility (*b* = 1.24, *p* = .380, *d* = 0.04), confidence in the average parameter estimate (*b* = 0.49, *p* = .749, *d* = 0.01), or ratings of bias (*b* = 2.13, *p* = .192, *d* = 0.06).     
As hypothesized, inconsistent crowd estimates lead to lower posterior beliefs, credibility of the results, and confidence in the average parameter estimate, and higher ratings of bias and error, than a single estimate. Unexpectedly, consistent crowd estimates lead to lower posterior beliefs and higher ratings of error than a single estimate. On all outcomes, consuming consistent crowd estimates performs worse or no better than consuming a single estimate. However, it is worth noting that there is a positive difference between prior and posterior beliefs within the consistent crowd condition ($M_d$ = 4.75, *d* = 0.19, *p* < .001; see Figure 2). This finding clarifies that a consistent crowd may sway observers --- however, relative to the conventional reporting of non-crowdsourced estimates, crowd estimates do not seem to improve lay perceptions of scientific findings.

>>>>

We sampled 1,498 participants from Prolific (native English speakers from the US and UK, over 18 years old). As preregistered, we regressed all outcomes on prior beliefs and condition (with the single estimate condition as the reference category). Compared to the single estimate condition, inconsistent crowd estimates decreased posterior beliefs (*b* = -22.80, *p* < .001, Cohen's *d* = -0.72), ratings of credibility (*b* = -6.54, *p* < .001, *d* = -0.21), and confidence in the parameter estimate (*b* = -9.09, *p* < .001, *d* = -0.26), and increased ratings of bias (*b* = 6.68, *p* < .001, *d* = 0.18), error (*b* = 5.89, *p* < .001, *d* = 0.18), and discretion (*b* = 9.22, *p* < .001, *d* = 0.29). Compared to the single estimate condition, consistent crowd estimates decreased posterior beliefs (*b* = -5.71, *p* < .001, *d* = -0.18) and increased ratings of error (*b* = 3.59, *p* = .016, *d* = 0.11) and discretion (*b* = 5.14, *p* < .001, *d* = 0.16); we found no significant effects on ratings of credibility (*b* = 1.24, *p* = .380, *d* = 0.04), confidence in the average parameter estimate (*b* = 0.49, *p* = .749, *d* = 0.01), or ratings of bias (*b* = 2.13, *p* = .192, *d* = 0.06).     
As hypothesized, inconsistent crowd estimates lead to lower posterior beliefs, credibility of the results, and confidence in the average parameter estimate, and higher ratings of bias and error, than a single estimate. Unexpectedly, consistent crowd estimates lead to lower posterior beliefs and higher ratings of error than a single estimate. On all outcomes, consuming consistent crowd estimates performs worse or no better than consuming a single estimate. However, it is worth noting that there is a positive difference between prior and posterior beliefs within the consistent crowd condition ($M_d$ = 4.75, *p* < .001, *d* = 0.19; see Figure 2). This finding clarifies that a consistent crowd may sway observers --- however, relative to the conventional reporting of non-crowdsourced estimates, crowd estimates do not seem to improve lay perceptions of scientific findings.

### 2021-10-16 poster/GSPA_Poster.Rmd
last changes prior to submission GSPA + ORA

Every day, important scientific findings are rejected at large. From man-made climate change to the safety and efficacy of Covid-19 vaccinations, **science skepticism** has run rampant among lay consumers in modern society (Hornsey & Fielding, 2017). To **increase public faith in science**, some have proposed the use of **crowd science** (Silberzahn et al., 2018; Uhlmann et al., 2019).  
 $~~~~~$ We explore whether high consensus and low variance in parameter estimates among a **crowd** (the **consistent crowd** condition) can improve **lay perceptions of scientific findings**. In line with **social norm theory** (Miller & Prentice, 2016), we expect that **observing consensus among a crowd** will -- compared to the conclusion of a single scientist (the **single estimate** condition) -- increase conformity in opinion. Drawing from work on **intuitive statistics** (Gigerenzer & Murray, 2015), we also expect laypeople to intuitively follow the logic of the **wisdom of crowds**: the ability of an **aggregate of estimates** (rather than a single estimate) to **reduce noise** stemming from individual bias or error (Schweinsberg et al., 2021).  
  $~~~~~$ In contrast, when crowd estimates show low consensus and high variance (the **inconsistent crowd** condition), we predict that observers will be less swayed and more likely to **attribute** the findings to **bias** and **error**. In addition, due to the difficulty of lay reasoning about variation (Ben-Zvi & Garfield, 1999), we predict an **aversion to variability**: i.e., we expect that observing variable estimates will decrease lay **confidence** in the precise average parameter estimate in both crowd conditions.

@@

**Figure 1**: *Estimates of differences with the single estimate condition*
```{r, echo = FALSE, out.width='93%', fig.retina=3, fig.align='left'}
knitr::include_graphics("Figure1.jpg")
```
**In line with our hypotheses**, lay consumers of **inconsistent crowd estimates** (vs. a single estimate)...  
`r icon_style(icons::fontawesome("arrow-circle-down"), scale = 1.5, fill = "red")` $~~$ Have lower posterior beliefs about the reported phenomenon  
`r icon_style(icons::fontawesome("arrow-circle-down"), scale = 1.5, fill = "red")` $~~$ Find the results less credible  
`r icon_style(icons::fontawesome("arrow-circle-down"), scale = 1.5, fill = "red")` $~~$ Have less confidence in the average estimate of 5%  
`r icon_style(icons::fontawesome("arrow-circle-up"), scale = 1.5, fill = "green")` $~~$ Are more likely to attribute the average estimate (5%) to bias   
`r icon_style(icons::fontawesome("arrow-circle-up"), scale = 1.5, fill = "green")` $~~$ Are more likely to attribute the average estimate (5%) to error

@@

**Future Directions**  
`r icon_style(icons::fontawesome("question-circle"), scale = 1.5, fill = "#035AA6")` $~~$ Does **variability aversion** explain the findings?  
`r icon_style(icons::fontawesome("flask"), scale = 1.5, fill = "#035AA6")` $~~~$ Perceptions of **scientists**  
`r icon_style(icons::fontawesome("comments"), scale = 1.5, fill = "#035AA6")` $~~$ **Science communication** and **communicating uncertainty**

>>>>

Every day, important scientific findings are rejected at large. From man-made climate change to the safety and efficacy of Covid-19 vaccinations, **science skepticism** has run rampant among lay consumers in modern society (Hornsey & Fielding, 2017). To **increase public faith in science**, some have proposed the use of **crowd science** (Silberzahn et al., 2018; Uhlmann et al., 2019).  
 $~~~~~$ We explore the effects of scientific findings emerging from a **crowd** of researchers (vs. a typical research collaboration) on **lay perceptions of scientific findings**. In line with **social norm theory** (Miller & Prentice, 2016), we expect that observing **consensus** among a crowd (the **consistent crowd** condition) will -- compared to the conclusion of a single scientist (the **single estimate** condition) -- increase conformity in opinion. Drawing from work on **intuitive statistics** (Gigerenzer & Murray, 2015), we also expect laypeople to intuitively accord to logic of the **wisdom of crowds**: the ability of an **aggregate of multiple estimates** (rather than a single estimate) to **reduce noise** stemming from individual bias or error (Schweinsberg et al., 2021).  
  $~~~~~$ In contrast, when crowd estimates show low consensus and high variance (the **inconsistent crowd** condition), we predict that observers will be less swayed and more likely to **attribute** the findings to **bias** and **error**. In addition, due to the difficulty of lay reasoning about variation (Ben-Zvi & Garfield, 1999), we predict an **aversion to variability**: i.e., we expect that observing variable estimates will decrease lay **confidence** in the precise average parameter estimate in both crowd conditions.

@@

**Figure 1**: *Estimates of differences with the single estimate condition*
```{r, echo = FALSE, out.width='93%', fig.retina=3, fig.align='left'}
knitr::include_graphics("Figure1c.jpg")
```
**In line with our hypotheses**, lay consumers of **inconsistent crowd estimates** (vs. a single estimate)...  
`r icon_style(icons::fontawesome("arrow-circle-down"), scale = 1.5, fill = "red")` $~~$ Have lower posterior beliefs about the reported phenomenon  
`r icon_style(icons::fontawesome("arrow-circle-down"), scale = 1.5, fill = "red")` $~~$ Find the results less credible  
`r icon_style(icons::fontawesome("arrow-circle-down"), scale = 1.5, fill = "red")` $~~$ Have less confidence in the average estimate of 5%  
`r icon_style(icons::fontawesome("arrow-circle-up"), scale = 1.5, fill = "green")` $~~$ Are more likely to attribute the average estimate (5%) to bias   
`r icon_style(icons::fontawesome("arrow-circle-up"), scale = 1.5, fill = "green")` $~~$ Are more likely to attribute the average estimate (5%) to error

@@

**Future directions**  
`r icon_style(icons::fontawesome("question-circle"), scale = 1.5, fill = "#035AA6")` $~~$ Does **variability aversion** explain the findings?  
`r icon_style(icons::fontawesome("flask"), scale = 1.5, fill = "#035AA6")` $~~~$ Perceptions of **scientists**  
`r icon_style(icons::fontawesome("comments"), scale = 1.5, fill = "#035AA6")` $~~$ **Science communication** and **communicating uncertainty**

### 2021-10-16 poster/ORAprojectsummary.Rmd
last changes prior to submission GSPA + ORA

---
title             : "Lay Perceptions of Scientific Findings: The Risks of Variability and Lack of Consensus"
shorttitle        : "Lay Perceptions of Scientific Findings"

author:
  - name          : "Insert first author name here"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
    address       : ""
    email         : ""
  - name          : "Insert second author name here"
    affiliation   : "1"

@@

abstract: |
  Crowd science is on the rise in behavioral science and aims to increase the credibility of scientific research. Whether it does, however, remains untested. We run an experiment to study whether crowd science improves lay perceptions of scientific findings. We focus on crowdsourced data analysis: a crowd of scientists independently analyzes the same dataset to estimate and report on a parameter of interest.  
  We compare the effects of providing lay consumers (*N* = 1,498) with a single non-crowdsourced parameter estimate (the single-analyst condition) vs. multiple crowdsourced parameter estimates that (a) vary slightly and are all positive, leading to the same qualitative conclusion (the "multi-consistent" condition) or (b) vary widely and are of both signs, leading to differing qualitative conclusions (the "multi-inconsistent" condition). In the single-analyst condition, a single team of six researchers reports a 5% estimate; in the multi-consistent condition, six independent researchers report six low variance, high consensus estimates (2%, 4%, 5%, 5%, 6%, 8%; $M = 5\%, \sigma = 2\%$); in the multi-inconsistent condition, six independent researchers report six high variance, low consensus estimates (-6%, -2%, 5%, 5%, 12%, 16%; $M = 5\%, \sigma = 8.25\%$).  
  According to our preregistered hypotheses and in line with social norm theory, when laypeople observe multiple consistent (inconsistent) estimates from independent scientists rather than a single estimate, we expect an increase (decrease) in posterior beliefs and credibility of the results, and a decrease (increase) in ratings of bias and error. In addition, we expect that observing differing estimates decreases confidence in the precise average estimate in both multi-estimate conditions.  `r knitr::load_cache("results")`
  In line with our hypotheses, we find that lay consumers of multiple, inconsistent (high variance, low consensus) crowdsourced estimates have lower posterior beliefs, `r final_inconsistent`; find the results less credible, `r credibility_inconsistent`; have less confidence in the average effect size estimate, `r confidence_inconsistent`; and believe the results are more likely to stem from bias, `r bias_inconsistent`, and error, `r error_inconsistent`, than consumers of a single, non-crowdsourced estimate. Contrary to our hypotheses, we do not find that consuming multiple, consistent (low variance, high consensus) crowdsourced estimates improves lay perceptions of scientific findings: instead, to our surprise, it leads to lower posterior beliefs, `r final`, and higher ratings of error, `r error`, than a single, non-crowdsourced estimate.

keywords          : "Crowd science, Many analysts, Multi-analyst, Variability, Research credibility"
wordcount         : "Abstract: 383, Project summary: x"

@@

The credibility of scientific research is in doubt, among lay consumer [@hornsey2017] and scientist [@pashler2012] alike. Several tools have been proposed to combat this "crisis of confidence" (Ibid., p. 528). One such tool is the crowd science approach: "the organization of scientific research in open and collaborative projects"  [@franzoni2014, p. 1]. We focus on crowdsourced data analysis, also known as the 'many analysts' or 'multi-analyst' approach: a crowd of scientists independently analyzes the same dataset to estimate and report on a parameter of interest.  
According to some of its proponents, crowdsourced data analysis should increase the credibility of scientific findings to lay consumers. In the preregistration of an ongoing many-analysts study, Arbon et al. [-@arbon2019] justify their use of the novel approach by referring to the need for "the public to have faith in the conclusions of scientists". In a preprint of a conducted multi-analyst study, Breznau et al. [-@breznau2021, p. 3] argue that independent data analyses with converging findings should offer "consumers of scientific findings assurance that they are not arbitrary flukes". In a commentary on the many analysts approach, Breznau [-@breznau2021a, p. 311], offers the hopeful view that crowdsourcing is a new way to "increase public, private, and government views of social science" and to "increase credibility for political and social research---in both sample populations and among the researchers themselves." Does crowd science meet this promise -- to improve lay perceptions of scientific findings -- in reality?  
We explore the effects of scientific findings emerging from a crowd of researchers (vs. a typical research collaboration) on lay consumers' posterior beliefs, perceptions of credibility, confidence in an aggregate effect size estimate, and ratings of bias and error. We compare the effects of providing lay consumers with a single parameter estimate (the single-analyst condition) vs. multiple crowdsourced parameter estimates that (a) vary slightly and are all positive, leading to the same qualitative conclusion (the "multi-consistent" condition) or (b) vary widely and are of both signs, leading to differing qualitative conclusions (the "multi-inconsistent" condition).    
In line with social norm theory [@miller2016], we expect an individual who observes a crowd of scientists reach identical (differing) qualitative conclusions to be more (less) likely to conform in opinion, and be more (less) swayed by the scientific findings, than an individual who observes the conclusion of a single scientist. However, regardless of the presence or absence of consensus in conclusions, we expect that observing (slightly to widely) varying estimates decreases lay confidence in the precise average estimate. In sum, our preregistered hypotheses are as follows: when laypeople observe multiple consistent (inconsistent) estimates from a crowd of independent scientists, we expect -- compared to a single, non-crowdsourced estimate -- higher (lower) posterior beliefs and credibility of the results, lower (lower) confidence in the precise average parameter estimate, and lower (greater) ratings of bias and error.

# Methods
All relevant materials -- including the preregistration, survey materials, data, and code -- are openly available on the OSF at [insert link]. We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. We sampled UK- and US-based participants from the Prolific participant pool (native English speakers, over 18 years old). We excluded participants who failed an attention check or attempted to take the survey more than once. Our target sample size was 1,500 participants after exclusions. 
After a brief introduction to a research question ("Do religious people report higher well-being?"), participants were asked to report their prior beliefs ("How likely do you think it is that people who are more religious generally report higher well-being?") on a slider from 0% (not likely at all) to 100% (extremely likely).  
After reporting their prior beliefs, participants were randomly allocated to one of three experimental conditions in which they learned about the approach and findings of a scientific study. In the single-analyst condition, a single team of six researchers reports a 5% increase in well-being among religious people; in the multi-consistent condition, six independent researchers report six  consistent estimates (2%, 4%, 5%, 5%, 6%, and 8%, respectively) that average to 5% (SD = 2); and in the multi-inconsistent condition, six independent researchers report six inconsistent estimates (-6%, -2%, 5%, 5%, 12%, and 16%, respectively) that average to 5% (SD = 8.25).     
Afterwards, participants rated (1) their posterior beliefs, (2) the credibility of the results, (3) their confidence in the effect size estimate, and how likely it is that the estimate was influenced by (4) bias, (5) error, and (6) researcher degrees of freedom. All questions were answered on a slider from 0% (not likely/credible/confident at all) to 100% (extremely likely/credible/confident).

## Data analysis
For all six measures, we run linear regression models with condition as the independent variable (with the single-analyst condition as the reference category) and prior beliefs as a covariate. All hypotheses, statistical models, and code were preregistered. We did not preregister any hypotheses for the last measure; the findings concerning the impact of experimental condition on ratings of researcher degrees of freedom are exploratory, and should be treated as such. Because we test five separate hypotheses using two comparisons each (one comparison of the single-analyst vs. the multi-consistent condition, and one comparison of the single-analyst vs. the multi-inconsistent condition), we use the Bonferroni method to correct for multiple comparisons. Thus, our preregistered threshold for statistical significance is $\displaystyle \alpha = \frac{.05}{2} = .025$.

@@

After two weeks of data collection, we recorded `r n_recorded` responses in Qualtrics. We excluded `r n_more_than_once` observations from participants who attempted to take the survey more than once, `r n_screened_out` participants who were screened out prior to starting the survey or did not consent, and `r n_failed_attention` participants who failed the attention check. This left us with a total sample of `r nrow(df)` participants (`r ns_bycondition[1]`, `r ns_bycondition[2]`, and `r ns_bycondition[3]` in the single-analyst, multi-consistent, and multi-inconsistent condition, respectively).     
Figure 1 displays the main findings of our linear regression models, which compare each condition to the single-analyst condition and control for prior beliefs. For the multi-inconsistent condition, we found significantly lower (a) posterior beliefs, `r results(Final_beliefs, "Inconsistent")`; (b) ratings of credibility, `r results(Credibility, "Inconsistent")`; and (c) confidence in the effect size estimate, `r results(Confidence, "Inconsistent")`; and significantly higher (d) ratings of bias, `r results(Bias, "Inconsistent")`; (e) error, `r results(Error, "Inconsistent")`; and (f) discretion, `r results(Discretion, "Inconsistent")`.  
For the multi-consistent condition, we found significantly lower (a) posterior beliefs, `r results(Final_beliefs)`; significantly higher (b) ratings of error, `r results(Error)`; and (c) discretion, `r results(Discretion)`, and no significant effects on (d) ratings of credibility, `r results(Credibility)`, (e) confidence in the effect size estimate, `r results(Confidence)`, or (f) ratings of bias, `r results(Bias)`.     
**[Insert Figure 1 here]**  
In line with our hypotheses, lay consumers of multi-analyst studies with inconsistent results (compared to single-analyst studies) have lower posterior beliefs, find the results less credible, have less confidence in the average effect size estimate, and believe the results are more likely to stem from bias and error. Contrary to our hypotheses, we do not find that lay consumers of multi-analyst studies with consistent results (compared to single-analyst studies) have higher posterior beliefs, find the results more credible, have less confidence in the average effect size estimate, and believe the results are less likely to stem from bias and error: instead, they report significantly lower posterior beliefs and are more likely to believe the results stem from error (we did not find significant effects on ratings of credibility, confidence, or bias). Figure 2 further clarifies the sway of multi-analyst vs. single-analyst studies, by displaying the distribution of prior and posterior beliefs across the three conditions.  
**[Insert Figure 2 here]**  
It is worth noting on the basis of Figure 2 and a post-hoc, paired *t*-test that, while multi-analyst studies with consistent results perform worse or no better than single-analyst studies on all measures, there is a significant, positive effect of the findings on posterior beliefs *within* the multi-consistent condition: i.e., beliefs in the research hypothesis are greater after reading the multi-consistent study results, `r t_apa`. This finding clarifies that multi-analyst studies are not necessarily bad in absolute terms --- however, when comparing to conventional, single-analyst scientific research, crowdsourced data analysis does not seem to provide an improvement in the sway and credibility of scientific research to lay consumers.

@@

From the proliferation of big team science and large-scale replication initiatives to preregistration and registered reports, several scientific fields have undergone significant reform with the well-intended goal of improving the reliability of scientific research. The multi-analyst approach has many worthy uses, from demonstrating the arbitrariness and impact of individual analytic choices to acknowledging the inherent variability of results and averaging across idiosyncratic analytic choices to obtain more accurate parameter estimates. However, as with any real-world intervention, scientific reform can have unintended consequences. Here, we focus on the effects of crowdsourcing data analysis, and find that the multi-analyst approach may have an unintended consequence.  
While partly instituted with the goal of improving the credibility of scientific research, lay consumers appear to resist the variability and lack of consensus that often comes with multi-analyst research. To our surprise, even when results generated by independent analysts are largely consistent, lay consumers are less likely to believe in the reported phenomenon and more likely to think that the findings stem from error and researcher degrees of freedom.

@@

## Figure 1
*Estimates relative to the single-analyst condition*
\noindent
```{r fig-1, echo=FALSE}
knitr::include_graphics("Figure1.jpg", dpi = 1200)
```
*Note.* Figure 1 displays coefficient estimates (and 95% confidence intervals) of posterior beliefs, credibility, confidence, bias, error, and discretion in the two multi-analyst conditions, compared to the single-analyst condition (and controlling for prior beliefs). Green/red/gray error bars indicate positive/negative/insignificant findings, respectively.

@@

*Individual data points, quartiles, and distributions of prior and posterior beliefs in the single-analyst, multi-consistent, and multi-inconsistent conditions*  
\noindent
```{r raincloud, echo=FALSE}
knitr::include_graphics("Figure2.jpg", dpi = 1200)
```
*Note. * Prior beliefs are displayed in blue; posterior beliefs are displayed in orange. The respective boxes display the lower quartiles, medians, and upper quartiles of prior and posterior beliefs by condition.

>>>>

---
title             : "Lay Perceptions of Scientific Findings: Swayed by the Crowd?"
shorttitle        : "Lay Perceptions of Scientific Findings"

author:
  - name          : "First author name"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
    address       : ""
    email         : ""
  - name          : "Second author name"
    affiliation   : "1"

@@

abstract: |
  Crowd science is on the rise in behavioral science and aims to increase the credibility of scientific research. Whether it does, however, remains untested. We run an experiment to study whether crowd science improves lay perceptions of scientific findings. We focus on crowdsourced data analysis: a crowd of scientists independently analyzes the same dataset to estimate and report on a parameter of interest.  
  We compare the effects of providing lay consumers (*N* = 1,498) with a single parameter estimate (the single estimate condition) vs. multiple parameter estimates from a crowd that (a) vary slightly and are all positive, leading to the same qualitative conclusion (the "consistent crowd" condition) or (b) vary widely and are of both signs, leading to differing qualitative conclusions (the "inconsistent crowd" condition). In the single estimate condition, a single team of six researchers reports a 5% estimate; in the consistent crowd condition, six independent researchers report six low variance, high consensus estimates (2%, 4%, 5%, 5%, 6%, 8%; $M = 5\%, \sigma = 2\%$); in the inconsistent crowd condition, six independent researchers report six high variance, low consensus estimates (-6%, -2%, 5%, 5%, 12%, 16%; $M = 5\%, \sigma = 8.25\%$). We draw from theories on social norms and numerical cognition to hypothesize that consistent (inconsistent) crowd estimates will lead to higher (lower) posterior beliefs and credibility of the results, lower confidence in the precise average parameter estimate, and lower (higher) ratings of bias and error, than a single estimate. `r knitr::load_cache("results")`  
  In line with our preregistered hypotheses, we find that lay consumers of inconsistent (high variance, low consensus) crowd estimates have lower posterior beliefs, `r final_inconsistent`; find the results less credible, `r credibility_inconsistent`; have less confidence in the average effect size estimate, `r confidence_inconsistent`; and believe the results are more likely to stem from bias, `r bias_inconsistent`, and error, `r error_inconsistent`, than consumers of a single estimate. Contrary to our hypotheses, we do not find that consuming consistent (low variance, high consensus) crowd estimates improves lay perceptions of scientific findings: instead, to our surprise, it leads to lower posterior beliefs, `r final`, and higher ratings of error, `r error`, than a single estimate. In the future, it is important for crowd scientists to consider how -- e.g., by drawing from research on science communication -- to tackle science skepticism and effectively communicate variable crowd estimates to lay consumers.

keywords          : "Crowd science, Science skepticism, Variability, Numerical cognition"
wordcount         : "Abstract: 400, Project summary: 965"

@@

Every day, important scientific findings are rejected at large. From man-made climate change to the safety and efficacy of Covid-19 vaccinations, science skepticism has run rampant among lay consumers in modern society [@hornsey2017]. To increase public faith in science, some have proposed the use of crowd science [@silberzahn2018; @uhlmann2019]. We explore the effects of scientific findings emerging from a crowd of researchers (vs. a typical research collaboration) on lay perceptions of scientific findings. In line with social norms theory [@miller2016], we expect that observing consensus among a crowd (the consistent crowd condition) will -- compared to the conclusion of a single scientist (the single estimate condition) -- increase conformity in opinion. Drawing from work on intuitive statistics [@gigerenzer2015a], we also expect laypeople to intuitively accord to the logic of "the wisdom of crowds": the ability of an aggregate of estimates (rather than a single estimate) to reduce noise stemming from individual bias or error [@schweinsberg2021].  
In contrast, when crowd estimates show low consensus and high variance (the inconsistent crowd condition), we predict that observers will be less swayed and more likely to attribute the findings to bias and error. In addition, due to the difficulty of lay reasoning about variation [@garfield2004], we predict an aversion to variability: i.e., we expect that observing variable estimates will decrease lay confidence in the precise average parameter estimate in both crowd conditions. In sum, our preregistered hypotheses are as follows: when laypeople observe multiple consistent (inconsistent) estimates from a crowd of independent scientists, we expect -- compared to a single estimate and controlling for prior beliefs -- higher (lower) posterior beliefs and credibility of the results, lower confidence in the precise average parameter estimate, and lower (greater) ratings of bias and error.

# Methods
We report how we determined our sample size, all data exclusions, all manipulations, and all measures. We sampled UK- and US-based participants from the Prolific participant pool (native English speakers, over 18 years old). There was no prior work to base an excepted effect size on; due to our expectation of relatively small effect sizes and monetary constraints, we decided on an upper limit of 1,500 participants after exclusions. After participants report their prior beliefs about a research question ("Do religious people report higher well-being?"), we randomly allocate them to one of three conditions. In the single estimate condition, a single team of six researchers reports a 5% increase in well-being among religious people; in the consistent crowd condition, six independent researchers report six low variance, high consensus estimates (2%, 4%, 5%, 5%, 6%, 8%; $M = 5\%, \sigma = 2\%$); in the inconsistent crowd condition, six independent researchers report six high variance, low consensus estimates (-6%, -2%, 5%, 5%, 12%, 16%; $M = 5\%, \sigma = 8.25\%$). Afterwards, participants rate -- on a slider from 0% to 100% -- their posterior beliefs, the credibility of the results, their confidence in the aggregate 5% estimate, and the likelihood of researcher bias, error, and discretion (i.e., idiosyncratic choices). We did not preregister any hypotheses for the last measure; rated discretion is thus an exploratory measure.

@@

We recorded 2,019 responses; after excluding `r n_more_than_once` observations from participants who attempted to take the survey more than once, `r n_screened_out` participants who were screened out prior to starting the survey or did not consent, and `r n_failed_attention` participants who failed the attention check, our total sample size was 1,498 participants. As preregistered, we regressed all outcomes on prior beliefs and condition (with the single estimate condition as the reference category). Figure 1 displays the main findings. Compared to the single estimate condition, inconsistent crowd estimates decreased posterior beliefs (*b* = -22.80, *p* < .001, Cohen's *d* = -0.72), ratings of credibility (*b* = -6.54, *p* < .001, *d* = -0.21), and confidence in the parameter estimate (*b* = -9.09, *p* < .001, *d* = -0.26), and increased ratings of bias (*b* = 6.68, *p* < .001, *d* = 0.18), error (*b* = 5.89, *p* < .001, *d* = 0.18), and discretion (*b* = 9.22, *p* < .001, *d* = 0.29). Compared to the single estimate condition, consistent crowd estimates decreased posterior beliefs (*b* = -5.71, *p* < .001, *d* = -0.18) and increased ratings of error (*b* = 3.59, *p* = .016, *d* = 0.11) and discretion (*b* = 5.14, *p* < .001, *d* = 0.16); we found no significant effects on ratings of credibility (*b* = 1.24, *p* = .380, *d* = 0.04), confidence in the average parameter estimate (*b* = 0.49, *p* = .749, *d* = 0.01), or ratings of bias (*b* = 2.13, *p* = .192, *d* = 0.06). As hypothesized, inconsistent crowd estimates lead to lower posterior beliefs, credibility of the results, and confidence in the average parameter estimate, and higher ratings of bias and error, than a single estimate. Unexpectedly, consistent crowd estimates lead to lower posterior beliefs and higher ratings of error than a single estimate. However, it is worth noting that there is a positive difference between prior and posterior beliefs within the consistent crowd condition ($M_d$ = 4.75, *p* < .001, *d* = 0.19; see Figure 2). This finding clarifies that a consistent crowd may sway observers --- however, compared to a conventional, non-crowdsourced estimate, crowd estimates do not seem to improve lay perceptions of scientific findings.

@@

Although crowd science aims to improve the credibility of scientific research, we find that lay consumers resist the variability and lack of consensus that often reflect its reality. To our surprise, even when results generated by a crowd are highly consistent, we find no improvement in lay perceptions. Further research is needed to consider the possible mechanisms behind this preliminary finding. In the future, it is important for crowd scientists to consider how to tackle science skepticism and effectively communicate variable estimates.

@@

## Figure 1
*Estimates of differences with the single estimate condition*
\noindent
```{r fig-1, echo=FALSE}
knitr::include_graphics("Figure1c.jpg", dpi = 1200)
```
*Note.* Figure 1 displays coefficient estimates (and 95% confidence intervals) of posterior beliefs, credibility, confidence, bias, error, and discretion in the two crowd conditions, compared to the single estimate condition (and controlling for prior beliefs). Green/red/gray error bars indicate positive/negative/insignificant findings, respectively.

@@

*Distribution of prior and posterior beliefs by condition*  
\noindent
```{r raincloud, echo=FALSE}
knitr::include_graphics("Figure2b.jpg", dpi = 1200)
```
*Note. * Prior beliefs are displayed in blue; posterior beliefs are displayed in orange. The respective boxes display the lower quartiles, medians, and upper quartiles of prior and posterior beliefs by condition.
